---
title: "R - Data Visualization of a Pharmaceutical Company"
author: "ABC"
date: "2022-11-28"
---

The objective is to understand the sales trends for one of the leading pharmaceutical company

#Import the packages
```{r}
require(magrittr)
require(dplyr)
require(ggplot2)
require(tidyr)
```

#Import the data
```{r}
sales_data = read.csv("SalesData.csv")
```

#1.	Compare Sales by region for 2016 with 2015 using bar chart
```{r}
#1. summarise the data 
    a1 = sales_data %>% 
         group_by(Region) %>% 
         summarise(Tot_Sales2015 = sum(Sales2015) , Tot_Sales2016 = sum(Sales2016))
    
    #2. gather the data 
    a2 = gather(a1, key = Year, value = Total_Sales,2:3)
    
    #3. plot the graph 
    ggplot(a2,aes(x= Region,y= Total_Sales,fill = Year , label = Total_Sales)) + geom_bar(stat = "identity",position = "dodge") + 
      ggtitle("Comparision of Sales by Region") + geom_text(size = 4)
```

```{r}
#Sales in the year 2016 are more than sales in 2015 across aa the regions.
```
    
    
#2.	Pie charts for sales for each region in 2016
```{r}
      #1. summarise the data 
    b1 = sales_data %>% group_by(Region) %>% summarise(Tot_Sales2016 = sum(Sales2016))
    View(b1)
    
    #2. plot the graph -1 
        pie(b1$Tot_Sales2016,b1$Region)
    
    #3. plot the graph -2
        b1$percent = round(b1$Tot_Sales2016 / sum(b1$Tot_Sales2016) * 100,2)
        label1 = paste(b1$Region,":",b1$percent,"%")
        
        pie(b1$Tot_Sales2016,label1,col = c("lightskyblue","lightgreen","lightpink"),main = "Sales 2016")
        
```        
        

#3.	Compare sales of 2015 and 2016 with Region and Tiers
```{r}
        #1. summarise the data
        c1 = sales_data %>% group_by(Region,Tier) %>% summarise(Tot_Sales2015 = sum(Sales2015) , Tot_Sales2016 = sum(Sales2016))
        View(c1)
        
        #2. gather the data
        c2 = gather(c1, key = Year, value = Total_Sales,3:4)
        View(c2)
        
        #3. plot the graph 
        ggplot(c2,aes(x= Tier,y= Total_Sales,fill = Year )) + geom_bar(stat = "identity",position = "dodge") + 
        ggtitle("Comparision of Sales by Region and Tier") + facet_wrap(~ Region)
        
```


#4.	In East region, which state registered a decline in 2016 as compared to 2015?
```{r}
        #1. summarise the data
        d1 = sales_data %>% filter(Region == "East") %>% group_by(State) %>% summarise(Tot_Sales2015 = sum(Sales2015) , Tot_Sales2016 = sum(Sales2016))
        View(d1)
        
        #2. gather the data
        d2 = gather(d1, key = Year, value = Total_Sales,2:3)
        View(d2)
        
        #3. plot the graph 
        ggplot(d2,aes(x= State,y= Total_Sales,fill = Year )) + geom_bar(stat = "identity",position = "dodge") + 
          ggtitle(" For East region - state wise yearly comparison") 
```

```{r}
#In East region, NY state decline in 2016 as compared to 2015.
``` 
  
      
#5.	In all the High tier, which Division saw a decline in number of units sold in 2016 compared to 2015?
```{r}        
        #1. summarise the data
        e1 = sales_data %>% filter(Tier == "High") %>% group_by(Division) %>% summarise(Tot_Units2015 = sum(Units2015) , Tot_Units2016 = sum(Units2016))
        View(e1)
        
        #2. gather the data
        e2 = gather(e1, key = Year, value = Total_Units,2:3)
        View(e2)
        
        #3. plot the graph 
        ggplot(e2,aes(x= Division,y= Total_Units,fill = Year )) + geom_bar(stat = "identity",position = "dodge") + 
          ggtitle(" For High Tier - division wise yearly comparison") + theme(axis.text.x = element_text(angle = 90))
``` 

    
#6.	Create a new column Qtr -
```{r}
    
    sales_data$Qtr = ifelse(sales_data$Month == "Jan" | sales_data$Month == "Feb" | sales_data$Month == "Mar" ,"Q1",
                     ifelse(sales_data$Month == "Apr" | sales_data$Month == "May" | sales_data$Month == "Jun" ,"Q2",  
                     ifelse(sales_data$Month == "Jul" | sales_data$Month == "Aug" | sales_data$Month == "Sep" ,"Q3" ,"Q4")))    
    View(sales_data)    
```     

       
#7.	Compare Qtr wise sales in 2015 and 2016 in a bar plot
```{r}    
    #1. summarise the data
    f1 = sales_data %>% group_by(Qtr) %>% summarise(Tot_Sales2015 = sum(Sales2015) , Tot_Sales2016 = sum(Sales2016))
    View(f1)
    
    #2. gather the data
    f2 = gather(f1, key = Year, value = Total_Sales,2:3)
    View(f2)
    
    #3. plot the graph 
    ggplot(f2,aes(x= Qtr,y= Total_Sales,fill = Year )) + geom_bar(stat = "identity",position = "dodge") + 
      ggtitle(" Qtr wise sales") 
```    
 
      
#8.	Determine the composition of Qtr wise sales in 2015 with regards to all the Tiers in a pie chart.
#(Draw 4 pie charts representing a Quarter for each Tier)
```{r}    
    #1. summarise the data
    g1_a = sales_data %>% filter(Qtr == "Q1") %>% group_by(Tier) %>% summarise(Tot_Sales2015 = sum(Sales2015))
    g1_a$percent = round(g1_a$Tot_Sales2015 / sum(g1_a$Tot_Sales2015) * 100,2)
    label1 = paste(g1_a$Tier,":",g1_a$percent,"%")
    
    g1_b = sales_data %>% filter(Qtr == "Q2") %>% group_by(Tier) %>% summarise(Tot_Sales2015 = sum(Sales2015))
    g1_b$percent = round(g1_b$Tot_Sales2015 / sum(g1_b$Tot_Sales2015) * 100,2)
    label2 = paste(g1_b$Tier,":",g1_b$percent,"%")
    
    g1_c = sales_data %>% filter(Qtr == "Q3") %>% group_by(Tier) %>% summarise(Tot_Sales2015 = sum(Sales2015))
    g1_c$percent = round(g1_c$Tot_Sales2015 / sum(g1_c$Tot_Sales2015) * 100,2)
    label3 = paste(g1_c$Tier,":",g1_c$percent,"%")
    
    g1_d = sales_data %>% filter(Qtr == "Q4") %>% group_by(Tier) %>% summarise(Tot_Sales2015 = sum(Sales2015))
    g1_d$percent = round(g1_d$Tot_Sales2015 / sum(g1_d$Tot_Sales2015) * 100,2)
    label4 = paste(g1_d$Tier,":",g1_d$percent,"%")
    
    
    #ploting them in one sheet
    par(mfrow = c(2,2))
    
    #2. Plot the graph
    pie(g1_a$Tot_Sales2015,label1,main = "Qtr 1")
    pie(g1_b$Tot_Sales2015,label2,main = "Qtr 2")
    pie(g1_c$Tot_Sales2015,label3,main = "Qtr 3")
    pie(g1_d$Tot_Sales2015,label4,main = "Qtr 4")
        
``` 
    
    